Motif roles provides unique information about species risk of extinction\\ \medskip Supplementary Information

### 2022-03-02 manuscript/reorder_main_text.tex
Updates from Overleaf

    %\subsection*{3 Global scale network properties are known to affect stability}
    Using Bayesian networks, researchers can explore how the structures of realistically-large networks shape their responses to disturbance.
    The global (whole-network) structural properties of ecological networks have implications for their functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002, Eklof2006}. 
    For example, more highly-connected networks are more resistant to secondary extinctions \citep{Dunne2002, Eklof2006} and a modular organization, where some groups of species are tightly interconnected with limited connections between groups, can limit the propagation of disturbances \citep{Krause2003, Teng2004}.

@@

    % Moving to network context
    As well as clarifying the relationships between network structure and the effect of bottom-up disturbance on particular species, motifs were also related to the mean probability of persistence across all species in a network.
    In fact, a network's motif profile had a stronger relationship to mean persistence compared to network size or connectance. 
    This suggests that network motif profiles may be more useful than global network properties when predicting mean extinction risk of species in a network.
    Connectance did, however, affect which motifs were most common within a network. 
    Highly-connected networks tended to contain higher frequencies of the omnivory motif, which includes more links than the other three-species motifs considered here.
    Global-scale properties may thus  have stronger effects on network persistence by changing meso-scale properties than directly.

>>>>

    %\subsection*{3 Global scale network properties are known to affect stability}
    Using Bayesian networks, researchers can explore how the structures of realistically-large networks shape their responses to disturbance.
    The global (whole-network) structural properties of ecological networks have implications for their functioning \citep{Petchey2002}, stability \citep{Allesina2012} and robustness to disturbances \citep{Dunne2002, Eklof2006}. 
    For example, more highly-connected networks are more resistant to secondary extinctions \citep{Dunne2002, Eklof2006} and a modular organization, where some groups of species are tightly interconnected with limited connections between groups, can limit the propagation of disturbances \citep{Krause2003, Teng2004, Gilarranz2017}.

@@

    % Moving to network context
    As well as clarifying the relationships between network structure and the effect of bottom-up disturbance on particular species, motifs were also related to the mean probability of persistence across all species in a network.
    In fact, a network's motif profile had a stronger relationship to mean persistence compared to network size or connectance. 
    This suggests that network motif profiles may be more useful than global network properties when predicting mean extinction risk of species in a network.
    Connectance did, however, affect which motifs were most common within a network. 
    Highly-connected networks tended to contain higher frequencies of the omnivory motif, which includes more links than the other three-species motifs considered here.
    Global-scale properties may therefore indirectly affect network persistence by changing meso-scale properties, such as the distribution of motifs.

### 2022-03-02 manuscript/supplement_motifs.tex
Updates from Overleaf

            \begin{itemize}
                \item Contains Table S4
            \end{itemize}    
    \end{enumerate}    
\end{spacing}
\clearpage

@@

        \begin{equation}
            \rho_{in} \approx \tau_{i} + S_{n}C_{n} + N_n,
            \label{partic_deg}
        \end{equation}

@@

        To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
        As with the PERMANOVA relating motif participation to network properties, we fit the PERMANOVA using the R~\citep{R} function `vegan' from the package \emph{adonis}~\citep{adonis} and calculated significance based on 9999 permutations.

@@

>>>>

            \begin{itemize}
                \item Contains Table S4
            \end{itemize}

        \item Glossary    
    \end{enumerate}    
\end{spacing}
\clearpage

@@

        \begin{equation}
            \rho_{in} \approx \tau_{i} + S_{n}C_{n} + N_n,
            \label{partic_tl}
        \end{equation}

@@

        To test whether networks with similar motif profiles tend to have similar mean persistences among consumers, we fit a PERMANOVA relating Bray-Curtis dissimilarity in motif profiles to the mean persistence of all consumers in a network (averaged across all levels of disturbance to basal resources).
        As with the PERMANOVA relating motif participation to network properties, we fit the PERMANOVA using the R~\citep{R} function `vegan' from the package \emph{adonis} and calculated significance based on 9999 permutations.

@@

\section{Glossary}

 \begin{table}[h!]
 \label{glossary}
 \caption{Glossary of terms relating to motifs and Bayesian networks}
     \footnotesize{
 \begin{tabular}{l|l}
     Term & Definition \\
     \hline
     Motif & Set of $n$ interacting species. In this case, $n=3$ \\
     Network motif profile & Vector describing the frequency of each motif in the network. \\
     & Normalised by dividing counts of each motif by the total across all motifs.\\
     Motif participation role & Vector describing the frequency with which a focal species appears in each motif.\\
     & Normalised by dividing counts for each motif by the total across all motifs. \\
     Bayesian network & A directed acyclic graph, used to predict species' likelihood of persistence. \\
     Network persistence & The mean likelihood of consumers in a network not going extinct.\\
     Species persistence & The individual likelihood of a species not going extinct.\\
     In-degree & Number of prey species to a consumer species.\\
     Trophic level (STL) & The shortest food chain between the focal species and any basal species.\\
     Disturbance ($\pi_{disturbed}$) & Probability of extinction of a basal resource when extra disturbance is added. \\
     Baseline extinction &  When no disturbance is applied, $\pi_{base}$=0.1.\\
 \end{tabular}}
 \end{table}
